### 2017-02-09 docs/IEEE/paper.tex
Wrote part of the introduction. 1st draft. Added some stuff to RW.txt

\documentclass[pageno]{jpaper}

\newcommand{\pactedition}{26th}
\newcommand{\pactyear}{17}
\newcommand{\pactcompyear}{20\pactyear}

%replace XXX with the submission number you are given from the PACT submission site.
\newcommand{\pactsubmissionnumber}{XXX}

\usepackage[normalem]{ulem}

\begin{document}

\title{Task Mining in Irregular Programs.}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}
We need to write our abstract here.
\end{abstract}

\section{Introduction}
\label{sec:intro}

% CONTEXT
Applications are usually written sequentially, regardless of the availability of a parallel architecture. Sequential code is often very irregular, and it is difficult for the compiler to be precise about whether a piece of code is independent or not.

%- Applications are usually written sequentially regardless of the availability of a parallel architecture
%- Hence sequential code is often irregular
%- However, sometimes many paralellism opportunities are missed during a sequential execution

% PROBLEM
Extensive work has been done to decide whether loops are irregular or parallel. However, due to the conservatively trait of the static analyses, very often a
loop or a code fragment is marked as irregular and dependent when in reality it might run in parallel most of the time. Finding this parallelism opportunities in irregular sequential code is hard because the programmer has to bear many aspects of it such as dependences between array-like structures.
The OpenMP tasking model allows the programmer to point out regions in the sequential code that can be parallelized, and although some work has been done to generate OpenMP annotations automatically, the burden of finding code regions to be annotated still lies upon the programmer.

% SOLUTION
Our work focuses on finding irregular parallelism and automatic generating compiler directives to make this parallelism explicit.
We present a fully automatic technique based on parallelism speculation that finds parallelism in programs and correctly annotates them so to make this parallelism explicit.
It is a speculative technique that consists in the combination of static and dynamic analyses that increase the volume of tasks that can run in parallel.
The static analysis finds irregular loops (i.e., loops with loop-carried dependences) and then speculates whether these loops can be safely executed in parallel. These loops
are annotated with compiler directives, to point out the possibility of task-parallelization within that sequence of code. Then a runtime able to compile these directives shall run them
and create the tasks for the annotated code. If the code is independent, the task-parallelized code will yield a great increase in performance. 
However, if the code is indeed dependent and cannot be executed in parallel, the runtime will solve the dependences and execute the code sequentially. 
This ensures correctness and makes the technique safe.

% TESTS AND RESULTS
We have implemented this technique in the LLVM compiler infrastructure. We tested our technique on three large benchmarks (CASTOR, LG Graphical Application, ???) and compared it to the manually annotated version. We show that XX\% of the manually annotated regions are found by our technique. Our technique also finds XX\% more possible task regions than what is manually annotated. Regarding performance, the automatically annotated code can be up to X times faster than the manually annotated one.

%Fernando's first draft
//DRAFT
The key contribution of this paper is a fully automatic technique to find parallelism in programs, and to annotate them so to make this parallelism explicit.
This technique consists in the combination of static and a dynamic analyses that increase the volume of tasks that can run in parallel.
The static analysis finds independent code, which can be safely executed in parallel.
However, due to its conservative nature, the static analysis misses potential parallelization opportunities.
These tasks, which we cannot show to be independent due, for instance, to poor aliasing information, are solved at runtime by the data-flow engine.
The final product of this combination is a high-performant framework that
lets us run algorithms that are traditionally difficult to parallelize.

\section{Overview}
\label{sec:ovf}

\section{Solution}
\label{sec:sol}

\section{Evaluation}
\label{sec:eval}

\section{Related Work}
\label{sec:rw}

\section{Conclusion}
\label{sec:conc}

\bstctlcite{bstctl:etal, bstctl:nodash, bstctl:simpurl}
\bibliographystyle{IEEEtranS}
\bibliography{references}

\end{document}