### 2017-02-11 docs/IEEE/paper.tex
Improved introduction

\documentclass[pageno]{jpaper}

\newcommand{\pactedition}{26th}
\newcommand{\pactyear}{17}
\newcommand{\pactcompyear}{20\pactyear}

%replace XXX with the submission number you are given from the PACT submission site.
\newcommand{\pactsubmissionnumber}{XXX}

\usepackage[normalem]{ulem}

\begin{document}

\title{Task Mining in Irregular Programs.}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}
We need to write our abstract here.
\end{abstract}

\section{Introduction}
\label{sec:intro}

% CONTEXT
It is undeniable that parallel computing is nowadays the most pursued alternative by the developers 
when striving for more application performance \cite{hwu2014ahead}. Two main reasons stand behind 
the actual urge for parallel programming. First, programmers often code in simple parallel patterns 
\cite{pinto2015large}, which evidences their awareness and will for parallelism. Second, the ever 
increasing industry of multi-cores and many-cores architectures propitiates the development of software 
with such behaviour \cite{haugen2016performance, vandierendonck2013analysis}. And although extensive 
work has been done to scavenge parallelism in regular applications \cite{kulkarni2009much}, finding good 
parallel opportunities in irregular code remains a grueling task.

% PROBLEM
Finding parallelism in sequentially-written code is hard because the programmer has to bear many aspects 
such as dependences between arrays, synchronization, scalability and race conditions.
Besides, even if we rely on static analyses to infer parallel regions, the results are imprecise 
\cite{kulkarni2007optimistic}. Due to the conservative trait of the static analyses, a
loop is frequently marked as iteration-dependent when, in reality, it might run in parallel most of the time. 
Some dependences pointed by the compiler might never occur during runtime. 
The OpenMP task scheduling model allows the programmer to point out regions in the sequential code that 
can be parallelized \cite{openmp2009}, which provides an easy way of parallelizing without reimplementation. 
Work has been carried out to generate OpenMP annotations automatically given a set of instructions that define 
a region \cite{gleison2016, pingali2011tao, wanggenerating}. Yet, the burden of finding code regions to be 
annotated with task directives still lies upon the programmer.

% SOLUTION
%Our work focuses on finding irregular code  and automatic generating compiler directives to make the parallelism explicit.
We present a fully automatic technique that finds parallelism in irregular programs and correctly annotates 
them so to make this pattern explicit.
It is a speculative technique that consists in the combination of static and dynamic analyses that increase 
the volume of tasks that can run in parallel.
The static analysis finds irregular loops (i.e., loops with loop-carried dependences) and then speculates 
whether these loops can be safely executed in parallel or not. These loops
are annotated with compiler directives, to point out the possibility of task-parallelization within that sequence of code. 
Then a runtime able to compile these directives runs the tasks for the annotated code. 
During execution, the dependences are resolved by the runtime and the code is executed in parallel.

The key insight of the tool is that it does not rely on the programmer to find possible parallel sites. 
It does so automatically, and the runtime is solely responsible for unraveling the dependences between tasks. 
The final product of this combination is a high-performant framework that
lets us run algorithms that are traditionally difficult to parallelize.

% TESTS AND RESULTS
We have implemented this technique in the LLVM compiler infrastructure. We tested our technique on three large benchmarks (CASTOR, LG Graphical Application, ???) and compared it to the manually annotated version. We show that XX\% of the manually annotated regions are found by our technique. Our technique also finds XX\% more possible task regions than what is manually annotated. Regarding performance, the automatically annotated code can be up to X times faster than the manually annotated one.

%Fernando's first draft
%//DRAFT
%The key contribution of this paper is a fully automatic technique to find parallelism in programs, and to annotate them so to make this parallelism explicit.
%This technique consists in the combination of static and a dynamic analyses that increase the volume of tasks that can run in parallel.
%The static analysis finds independent code, which can be safely executed in parallel.
%However, due to its conservative nature, the static analysis misses potential parallelization opportunities.
%These tasks, which we cannot show to be independent due, for instance, to poor aliasing information, are solved at runtime by the data-flow engine.

\section{Overview}
\label{sec:ovf}

\section{Solution}
\label{sec:sol}

\section{Evaluation}
\label{sec:eval}

\section{Related Work}
\label{sec:rw}

\section{Conclusion}
\label{sec:conc}

\bstctlcite{bstctl:etal, bstctl:nodash, bstctl:simpurl}
\bibliographystyle{IEEEtranS}
\bibliography{references}

\end{document}