### 2018-06-06 docs/ACM/Rebuttal.txt
Adding rebuttal.

We thank the referees for all the time and expertise they have put into reviewing our work. We address individual comments below.

Review #124A

> The example in Fig-3 shows limitations of TaskMiner as currently implemented. The generated code uses the if clause on the OpenMP task construct to limit generation of tasks. However, the if clause leads to the generation of an undeferred task. While the task must be executed immediately, most of the overheads of OpenMP task generation are still implied. The implementation must create a new data environment for the task. The correct way to end the task generation requires the use of the final clause, which leads to tasks that are mergeable and thus can be merged tasks, for which a data environment does not need to be created.

We agree with the referee: 'final+mergeable' is better. Gcc did not support 'final' when we started the project (see, the last slide of https://openmp.org/wp-content/uploads/sc15-openmp-CT-MK-tasking.pdf). However, although not ideal, the 'if' clause is correct, and it brings runtime gains. We have moved into 'final' already (we use 'final' to limit recursion, and 'if' to implement the dynamic cost model). The fibonacci example in the paper finds the 40th fib number in:

'omp-final': real=7.68s/user=14.25s;
'omp-if': real=7.51s/user=37.48s;
'omp+nothing, aka no-optimization': real=1m46.57s/user=11m16.72s

> The authors do not discuss the restrictions of the depend clause on OpenMP Task constructs. "List items used in depend clauses of the same task or sibling tasks must indicate identical storage locations or disjoint storage locations." The implementation does not appear to consider this restriction so it is unclear if the automatically generated OpenMP tasking code is correct. This issue is significant as the proposed implementation may yield significantly higher parallelism.

We consider those restriction! We follow Section 2.13.9 of the standard. Even more, we only annotate a region if we can find symbolic limits for all the memory accesses used within it. We only annotate arrays if we can determine their sections. Whenever we annotate different sections within the same array, these sections are disjoint across tasks. Another restriction is that a region must be profitable (see Page10, Line34, Column1). Given these two restrictions, we found "annotable regions" in 63 out of 219 benchmarks in the test suite. Every benchmark contains sanity checks.

> The approach should generate taskloop and taskgroup constructs.

We agree, and we are working to accommodate task grouping.

> Cilk++ is not annotation based. It's inclusion in the list in the introduction is incorrect.

Ok.

> The default clause cannot be specified with the private parameter in C and C++.

We believe the referee refers to Page4, Line38, Column2, where we wrote: "default([shared/private])". That was bad writing. We use private and firstprivate, but never within default. We shall fix that. Notice that in Page8, Line12, Column1, we explain that 'shared' is the only default.

> The y-axis on Figure 10 needs a label. What does the graph show? I think it is speedup compared to the sequential version of the program.

Yes, the referee's intuition is correct. We shall fix it.

> Including "bel-ford" as one of the three programs that most benefits from your cost model is not convincing since it appears to suffer from slowdown compared to the sequential version. Why doesn't your cost model determine that using tasks is a bad idea?

Some of the programs that we annotate experiment slowdowns. That's like compiling with gcc-O3, for instance: it slows down some programs, while speeding up others. The problem with bellman-ford is not due to the cost model. This algorithm traverses a graph implemented as an array of arrays. Dependences hinder parallelism whenever two tasks receive the adjacent list of the same node. We experimented similar problems with other irregular benchmarks (see Page10, Line53, Column1).

Review #124B

> Comparing against manual parallel implementation makes sense, but I would appreciate to see alternative approaches.

We are comparing against gcc-O3, with vectorization enabled. The Swan benchmarks were taken from the Intel SPMD compiler (ISPC). We can compare against them. However, each benchmark is written in C and in SPMD. Only the latter is parallelized by ISPC (using vector instructions).

> In the evaluation section I would appreciate to see an evaluation of the cost model used in this work.

Figure 11 presented an evaluation in three benchmarks. We also would like to show numbers about some tuning of it. We did not do it due to lack of space.

Review #124C

> How the algorithm determines where spawned tasks will be synchronized?

We shall clarify it. We synchronize at the post dominator of each parallel region.

Review #124D

> The paper is well written, with clear motivation and easy to understand. The challenge with work like this is that while it provides a great tool with useful techniques, the chosen techniques by themselves are not very novel and therefore the insights generated may be limited. That said, the tool is quite valuable and the design and implementation of it will be useful to the community, especially when made available publicly. The paper will serve as a useful report of the workings and principles behind the tool but it is unclear whether the PACT audience will find it directly useful due to the limited novelty/insights.

We thank the referee for the kind words. We agree that we are not providing any fundamentally new technique. We closed our paper with this remark: "This methodology does not introduce any fundamentally new static analysis or code optimization; in this field, we claim no contribution. Instead, our contributions lay into the overall design that eventually emerged from a two-years long effort to combine existing compilation techniques towards the goal of profiting from the powerful runtime system that OpenMP’s task parallelism brings forward". Thus, we would like to believe that PACT is the right community for our work, as it fits well the topic "Compilers and tools for parallel computer systems", which is available in the call for papers.